---
title: "DataGroup"
output: html_document
date: "2025-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Mental Health and Social - Good and Excellent

```{r}
library(tidyverse)
data = read_csv("mental_health_and_technology_usage_2024.csv")

data_mental_good = data%>%
  filter(Mental_Health_Status == "Good")

Social_Good <- data_mental_good %>%
  mutate(Hour_Range = cut(Social_Media_Usage_Hours, 
                          breaks = seq(0, ceiling(max(Social_Media_Usage_Hours, na.rm = TRUE)), by = 1), 
                          right = FALSE, 
                          labels = paste0(seq(0, ceiling(max(Social_Media_Usage_Hours, na.rm = TRUE)) - 1), 
                                          "-", 
                                          seq(1, ceiling(max(Social_Media_Usage_Hours, na.rm = TRUE)))))) %>%
  group_by(Hour_Range) %>%
  summarise(num_people = n()) %>%
  ggplot(mapping = aes(x = Hour_Range, y = num_people, fill = "blue")) +
  geom_col(show.legend = FALSE) +
  labs(x = "Social Media Usage Hours", y = "Number of People", title = "Distribution of Social Media Usage With Good Mental Health") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # 旋转x轴标签，防止重叠



Social_Good




long_data <- data_mental_good %>%
  mutate(Social_Hours = floor(Social_Media_Usage_Hours),
         Gaming_Hours = floor(Gaming_Hours),
         Screen_Hours = floor(Screen_Time_Hours)) %>%
  pivot_longer(cols = c(Social_Hours, Gaming_Hours, Screen_Hours), 
               names_to = "Category", values_to = "Hours") %>%
  group_by(Category, Hours) %>%
  summarise(num_people = n(), .groups = "drop") %>%
  ggplot(aes(x = factor(Hours), y = num_people, fill = Category)) +  # 使用 Category 作为 fill
  geom_col(position = "dodge") +  # dodge 让不同类别的柱子并排显示
  labs(x = "Usage Hours", y = "Number of People", title = "Comparison of Different Usage Hours With Good Mental Health") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # 旋转x轴标签，防止重叠

print(long_data)

```
```{r}
data_mental_excellent = data%>%
  filter(Mental_Health_Status == "Poor")

Social_Excellent <- data_mental_excellent %>%
  mutate(Hour_Range = cut(Social_Media_Usage_Hours, 
                          breaks = seq(0, ceiling(max(Social_Media_Usage_Hours, na.rm = TRUE)), by = 1), 
                          right = FALSE, 
                          labels = paste0(seq(0, ceiling(max(Social_Media_Usage_Hours, na.rm = TRUE)) - 1), 
                                          "-", 
                                          seq(1, ceiling(max(Social_Media_Usage_Hours, na.rm = TRUE)))))) %>%
  group_by(Hour_Range) %>%
  summarise(num_people = n()) %>%
  ggplot(mapping = aes(x = Hour_Range, y = num_people, fill = "blue")) +
  geom_col(show.legend = FALSE) +
  labs(x = "Social Media Usage Hours", y = "Number of People", title = "Distribution of Social Media Usage With Excellent Mental Health") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # 旋转x轴标签，防止重叠



Social_Excellent

long_data <- data_mental_excellent %>%
  mutate(Social_Hours = floor(Social_Media_Usage_Hours),
         Gaming_Hours = floor(Gaming_Hours),
         Screen_Hours = floor(Screen_Time_Hours)) %>%
  pivot_longer(cols = c(Social_Hours, Gaming_Hours, Screen_Hours), 
               names_to = "Category", values_to = "Hours") %>%
  group_by(Category, Hours) %>%
  summarise(num_people = n(), .groups = "drop") %>%
  ggplot(aes(x = factor(Hours), y = num_people, fill = Category)) +  # 使用 Category 作为 fill
  geom_col(position = "dodge") +  # dodge 让不同类别的柱子并排显示
  labs(x = "Usage Hours", y = "Number of People", title = "Comparison of Different Usage Hours with Excellent mental health") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # 旋转x轴标签，防止重叠

print(long_data)


```

starting with some anova and predictions
```{r}
library(tidyverse)
data01 = data %>% 
  mutate(mental_scale_value = case_when(
    Mental_Health_Status == "Excellent" ~ 2,
    Mental_Health_Status == "Good" ~ 1,
    Mental_Health_Status == "Fair" ~ 0,
    Mental_Health_Status == "Poor" ~ -1
  ))%>%
  mutate(stress_level = case_when(
    Stress_Level == "High" ~ 1,
    Stress_Level == "Medium" ~ 0,
    Stress_Level == "Low" ~ -1
  ))%>%
  select(Age,Gender,Technology_Usage_Hours ,Social_Media_Usage_Hours, Gaming_Hours,stress_level,Screen_Time_Hours,Sleep_Hours,mental_scale_value ,Support_Systems_Access,Work_Environment_Impact)
str(data01)
```


```{r}
mod_mental = lm(mental_scale_value ~ ., data = data01)
summary(mod_mental)
```
```{r}

# 自动变量选择（前向、后向、逐步）
s = step(lm(mental_scale_value ~ ., data = data01), direction = "both")
```
```{r}
final = lm(mental_scale_value ~ Support_Systems_Access, data = data01)
final
```
We can see that the final didn't include stress and screen time.
Why did screen_time and stress_level not work?
Some possible reasons:

The relationship between screen_time/stress_level and mental health is weak, so they don’t significantly affect the model.

You might not have enough data to detect a significant relationship, even if one exists.

Multicollinearity: Variables like screen_time and tech_usage might be too similar, causing one to mask the effect of the other.

Measurement error: For example, if the mental_health scale is too narrow or lacks standardization.


```{r}
str(data)
model_interaction <- lm(mental_scale_value ~ stress_level * Screen_Time_Hours, data = data01)
summary(model_interaction)
```


stress level 和 mental_scale_value 之间没有显著关系


```{r}
data01$stress_group <- cut(data01$stress_level,
                           breaks = quantile(data$stress_level, probs = c(0, 0.33, 0.66, 1)),
                           labels = c("Low", "Medium", "High"),
                           include.lowest = TRUE)

aov_result <- aov(mental_scale_value ~ stress_group, data = data01)
summary(aov_result)

```

```
```{r}

# 用中位数划分心理健康：高于中位数 = "Good"，否则 = "Poor"
median_mental <- median(data01$mental_scale_value, na.rm = TRUE)

data01$mental_binary <- ifelse(data01$mental_scale_value > median_mental, "Good", "Poor")

# 把它转换为因子变量
data01$mental_binary <- factor(data01$mental_binary, levels = c("Poor", "Good"))

logit_model <- glm(mental_binary ~ stress_level + Screen_Time_Hours,
                   data = data01, family = binomial)

summary(logit_model)


```

```{r}
# 用模型做预测（返回概率）
data01$predicted_prob <- predict(logit_model, type = "response")

# 转成分类结果（设定阈值 0.5）
data01$predicted_class <- ifelse(data01$predicted_prob > 0.5, "Good", "Poor")
data01$predicted_class <- factor(data01$predicted_class, levels = c("Poor", "Good"))

# 混淆矩阵
table(Predicted = data01$predicted_class, Actual = data01$mental_binary)

# 准确率
mean(data01$predicted_class == data01$mental_binary)

```







